CrypTool Script \\ Mathematics and Cryptography

### 2002-06-28 script/de/chapters/ellipticcurves.tex
ec+nt updates

Setzt man den Trend, dass sich die
Leistung der verf"ugbaren Rechner im Schnitt alle 18 Monate verdoppelt (Gesetz von Moore\index{Gesetz von Moore}), in die Zukunft
fort, so kann man von einer Entwicklung der ben"otigten Schl"ussell"angen wie in der folgenden Graphik ausgehen
(Quelle: Arjen Lenstra und Eric Verheul: 
\href{http://cryptosavvy.com/table.htm}{\texttt{http://cryptosavvy.com/table.htm}}).

\vskip -40 pt
% -> Figur 1
\begin{figure}[h]
\begin{center}
\vspace{1.5cm}\includegraphics[scale=0.75]{figures/RSAKeylength}
\caption{Gegenüberstellung des Aufwands der Operationen Sign und Verify bei
  RSA und Elliptischen Kurven (Quelle: Dr. J.
  Merkle, Elliptic Curve Cryptography Workshop, 2001)} 
\end{center}
\end{figure}
\vskip -30 pt

@@

Bei der digitalen Signatur muss man differenzieren: f"ur die {\em Erstellung} einer digitalen Signatur ben"otigen
bei auf Elliptischen Kurven basierende Verfahren im Vergleich zu RSA nur gut ein
Zehntel des Rechenaufwandes (66 zu 515 Multiplikationen). Betrachtet man die f"ur eine {\em Verifikation} durchzuf"uhrenden
Rechenschritte, dreht sich dieses Bild jedoch zu Gunsten des RSA um. Der Grund liegt darin, dass es bei
Verwendung des RSA m"oglich ist, einen sehr kurzen "offentlichen Schl"ussel zu w"ahlen, solange der
private Schl"ussel nur hinreichend lang ist.

% -> Figur 2
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.75]{figures/ThousandBitMultiplications}
\caption{Prognose f"ur die Entwicklung der als sicher betrachteten
  Schl"ussell"ange f"ur RSA und Elliptische Kurven\vspace{1ex}} 
\end{center}
\end{figure}
\vskip -20 pt

>>>>

Setzt man den Trend, dass sich die
Leistung der verf"ugbaren Rechner im Schnitt alle 18 Monate verdoppelt (Gesetz von Moore\index{Gesetz von Moore}), in die Zukunft
fort, so kann man von einer Entwicklung der ben"otigten Schl"ussell"angen
wie in Abbildung~\ref{RSAKeylength} ausgehen
(Quelle: Arjen Lenstra und Eric Verheul: 
\href{http://cryptosavvy.com/table.htm}{\texttt{http://cryptosavvy.com/table.htm}}).

\vskip -40 pt
% -> Figur 1
\begin{figure}[h]
\begin{center}
\vspace{1.5cm}\includegraphics[scale=0.75]{figures/RSAKeylength}
\caption{Gegenüberstellung des Aufwands der Operationen Sign und Verify bei
  RSA und Elliptischen Kurven} 
\label{RSAKeylength}
\end{center}
\end{figure}
\vskip -30 pt

@@

Bei der digitalen Signatur muss man differenzieren: f"ur die {\em
  Erstellung} einer digitalen Signatur ben"otigen bei auf Elliptischen
Kurven basierende Verfahren im Vergleich zu RSA nur gut ein Zehntel des
Rechenaufwandes (66 zu 515 Multiplikationen). Siehe hierzu
Abbildung~\ref{ThousandBitMultiplications} (Quelle: Dr. J.  Merkle,
Elliptic Curve Cryptography Workshop, 2001) Betrachtet man die f"ur eine
{\em Verifikation} durchzuf"uhrenden Rechenschritte, dreht sich dieses Bild
jedoch zu Gunsten des RSA um. Der Grund liegt darin, dass es bei Verwendung
des RSA m"oglich ist, einen sehr kurzen "offentlichen Schl"ussel zu
w"ahlen, solange der private Schl"ussel nur hinreichend lang ist.

% -> Figur 2
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.75]{figures/ThousandBitMultiplications}
\caption{Prognose f"ur die Entwicklung der als sicher betrachteten
  Schl"ussell"ange f"ur RSA und Elliptische Kurven\vspace{1ex}} 
\label{ThousandBitMultiplications}
\end{center}
\end{figure}
\vskip -20 pt

### 2002-06-28 script/en/chapters/ellipticcurves.tex
misc
ec+nt updates

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(Filipovics B. / Esslinger B. / Oyono R., April 2000, Update Dec. 2001)

@@

\section*{URLs}\addcontentsline{toc}{subsection}{URLs}

>>>>

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(Filipovics B. / Esslinger B. / Oyono R., April 2000, Update Dec. 2001,
June 2002)

\subsection{Elliptic curce cryptography --- a high-performance substitute for RSA?}

In the financial business secure and efficient data transfer is essential.
In particular, the RSA algorithm is used in many applications. Although the
security of RSA is beyond doubt, the evolution in computing power has
caused a growth in the necessary key length.  Today, 1024-bit RSA-keys are
standard, but the The GISA\index{GISA} (German Information Security Agency)
recommends the usage of 2048-bit keys from 2006 on (compare
section~\ref{SecurityRSA}).  The fact that most chips on smart cards cannot
process keys extending 1024 bit shows that there is a need for
alternatives. Elliptic curve cryptography (ECC) can be an such an
alternative in the field of asymmetric cryptography.

The efficiency of a cryptographic algorithm depends on the key length and
the calculation effort that is necessary to provide a prescribed level of
security. The major advantage of ECC compared to RSA is that it requires
much shorter key lengths. If we assume that the computing power increases
by Moore’s law (i.~e.\ it doubles every 18 months), then the evolution of
the key lengths for secure communication will be as
figure~\ref{ThousandBitMultiplications} (source: Arjen Lenstra und Eric
Verheul:
\href{http://cryptosavvy.com/table.htm}{\texttt{http://cryptosavvy.com/table.htm}}).

% -> Figure 1
\begin{figure}[h]
\begin{center}
\vspace{1.5cm}\includegraphics[scale=0.75]{figures/RSAKeylength}
\caption{Comparison of signing and verification time for RSA and Elliptic Curves} 
\label{RSAKeylength}
\end{center}
\end{figure}

In addition, a digital signature can be processed 10-times faster with ECC
than with RSA.  However, verification of a given signature is still more
efficient with RSA than with ECC. Refer to
figure~\ref{ThousandBitMultiplications} (source: Dr.~J.\ Merkle, Elliptic
Curve Cryptography Workshop, 2001) for a comparison.  The reason is that
RSA public keys can be chosen relatively small as long as the secret key is
sufficiently long.

% -> Figure 2
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.75]{figures/ThousandBitMultiplications}
\caption{Prognose of the key lengths to be regarded safe for RSA and
  Elliptic Curves\vspace{1ex}} 
\label{ThousandBitMultiplications}
\end{center}
\end{figure}

Nevertheless, thin clients like smart cards usually have to store the (long) secret key
and have to process a digital signature rather than to verify one. Therefore, there is
a clear advantage for ECC in terms of efficiency.
\par
\smallskip
Today, the major problem with ECC-implementations is the lack of standardization.
There is only one way to implement RSA, but there are many ways for ECC: One can work with
different sets of numbers, different (elliptic) curves described by up to 6 parameters,
and a variety of representations of the elements on the curve. Each choice has its
advantages and disadvantages, and one can certainly construct the most efficient for
each application. However, this causes problems in interoperability. But if all
ECC-tools should be able to communicate with each other, they will have to support
all different algorithms, which might put the advantage of efficient computation and
the need of less storage capacity to the contrary.

Therefore, international standardization organizations like IEEE (P1363),
ASC (ANSI X9.62, X9.63), ISO/IEC as well as major players like RSA labs or
Certicom have recently started standardization initiatives. While the IEEE
only describes the different implementations, the ASC has explicitly stated
10 elliptic curves and recommends their usage. The advantage of the ASC
approach is that one needs only a single byte to indicate which curve is
meant. However, it is not clear yet whether the ASC-curves will become a de
facto standard.

Although we see no need to replace RSA in any application today, one should
take the usage of ECC-based tools into consideration whenever a new system
is set up --- in particular, when the tool should be available beyond 2005.

@@

\section*{Web links}\addcontentsline{toc}{subsection}{Web links}

### 2002-06-28 script/en/chapters/numbertheory.tex
ec+nt updates

\newpage
%\hypertarget{Kap_3}
\section{Introduction to Elementary Number Theory with Examples}
(Bernhard Esslinger, besslinger@web.de, July 2001, Update Dec. 2001)

@@

\begin{center}
\fbox{\parbox{15cm}{
    \emph{Seneca\footnotemark:}\\
    The way of theory is long, it is short and effective by examples.
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  Lucius Annaeus, philosophical writer and poet, 4 B.~C. $-$ 65 A.~D.
}

@@

@@

\bibitem[Yan2000]{Yan2000} \index{Yan 2000} 
    Song Y. Yan, \\
    Number Theory for Computing, Springer, 2000.
\end{thebibliography}

@@

          \href{http://www.cryptool.com}{\texttt{http://www.cryptool.com}}
   \item Mathematica, \\
    Commercial mathematics package \\
          \href{http://www.wolfram.com}{\texttt{http://www.wolfram.com }}
   \item  LiDIA,\index{LiDIA} \\
      Extensive library containing number-theory functions and the LC interpreter \\
          \href{http://www.informatik.tu-darmstadt.de/TI/LiDIA}{\texttt{http://www.informatik.tu-darmstadt.de/TI/LiDIA}}
   \item BC, \\
     Interpreter with number-theory functions \\
          \href{http://www.maths.uq.edu.au/~krm/gnubc.html}{\texttt{http://www.maths.uq.edu.au/\~{}}krm/gnubc.html}
   \item Only after I had completed this article I did come across a website which interactively and 
     didactically uses elementary number theory to provide an extremely sophisticated description of the
     fundamental mathematical thought processes. It was created for a teaching project in the 11th grade 
     of the technical grammar school (unfortunately only available in German):\index{Muenchenbach} \\
         \href{http://www.djh-freeweb.de/~muenchenbach}{\texttt{http://www.djh-freeweb.de/\~{}muenchenbach}}
    \item GISA,\index{GISA} \\
          German Information Security Agency \\
          \href{http://www.bsi.bund.de}{\texttt{http://www.bsi.bund.de}} 
\end{enumerate}

@@

>>>>

\newpage
%\hypertarget{Kap_3}
\section{Introduction to Elementary Number Theory with Examples}
(Bernhard Esslinger, besslinger@web.de, July 2001, Update Dec. 2001, June 2002)

@@

\begin{center}
\fbox{\parbox{15cm}{
    \emph{Seneca\footnotemark:}\\
    The way of theory is long, it is short and effective by examples.
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  Lucius Annaeus Seneca, philosophical writer and poet, 4 B.~C. $-$ 65 A.~D.
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Considerations regarding the security of the RSA algorithm}\label{SecurityRSA}

Major parts of this chapter come from the article ``Vorz\"uge und Grenzen des
RSA-Verfahrens'' written by F.~Bourseau / D.~Fox / C.~Thiel, and published
in \emph{Datenschutz und Datensicherheit} (DuD), 26/2002, pp. 84$-$89.

There have always been discussions about the suitability of the RSA
algorithm for digital signatures and encryption, e.~g.\ after publications of
breakthroughs in factorization. Nevertheless the RSA algorithm has become a
de-facto standard since it was published more than 20 years ago.

The security of the RSA algorithm rests --- as with all cryptographic
methods --- on the following 4 central pillars:
\begin{itemize}
\item the complexity of the number theoretical problem on which the
  algorithm is based (here factorization of big numbers),
\item the election of fitting parameters (here the length of the module $N$),
\item the adequate usage of the algorithm and key generation and
\item the correct implementation of the algorithm.
\end{itemize}
Usage and key generation are well understood today. Implementation based on
long integer arithmetic is very easy.

%Therefor the fundamental characteristics of a special method are the first two points.
The following sections examine the RSA algorithm with respect to the first
two points.

\subsubsection{Complexity}

Successful decryption or forgery of a signature --- without knowing the
private key --- requires calculating the $e$-th root mod $n$.  The private
key, this is the multiplicative inverse of $e$ mod $J(n)$, can be easily
determined if $J(n)$ is known.  $J(n)$ again can be calculated from the
prime factors of $n$.  Breaking of RSA therefore cannot be more difficult
than factorization of the module $n$.

The best factorization method known today is a further development of the
General Number Field Sieve (GNFS), which was originally devised to
factor only numbers of a special form (like Fermat numbers).  The
complexity of solving the factorization problem with the GNFS is
asymptotically
$$
O(l) = e^{c \cdot (l \cdot \ln 2)^{1/3} \cdot  (\ln(l \cdot \ln(2))^{2/3} + o(l)}
$$
Please refer to:
\begin{list}{}{\setlength{\topsep}{-7 pt}}
\item[] A. Lenstra / H. Lenstra: \\ {\em The development of the Number
    Field Sieve.}  Lecture Notes in Mathematics 1554, Springer, New York
  1993
\item[] Robert D. Silverman: \\ {\em A Cost-Based Security Analysis of
    Symmetric and Asymmetric Key Lengths.}  In: RSA Laboratories Bulletin,
  No. 13, April 2000, S. 1-22
\end{list} \vskip +12 pt

This formula shows, that the factorization problem belongs to the class of
problems with sub-exponential time complexity (i.~e.\ time complexity grows
asymptotically not as fast as exponential functions like $e^l$ or $2^l$,
but stricly slower, e.~g.\ like $e^{\sqrt{l}}$).  This classification is
all that is currently known; it does not preclude the possiblity that the
factorization problem can be solved in polynomial time (s.
\ref{RSABernstein}).

$O(l)$ is the average number of processor steps depending on the bit length
$l$ of the number $n$ to be factored.  For the best currently known
factorization algorithm the constant $c = (64/9)^/173) = 1,923$.

The inverse proposition, that the RSA algorithm can be broken only by
factorization of $n$, is still not proven.  Most number theorists consider
the ``RSA problem'' and the factorization problem equivalent in terms of
time complexity.

Please refer to:
\begin{list}{}{\setlength{\topsep}{-7 pt}}
   \item[] A.J. Menezes / van Oorschot / P.C. Vanstone: \\
           {\em Handbook of Applied Cryptography}, CRC Press, 1996
\end{list}\vskip +12 pt

\subsubsection{Security Parameters}

The complexity is basically determined by the length $l$ of the module $n$.

In 1994 a 129-digit RSA module (428 bit), published in 1977, was factorized
by a distributed implementation of the Quadratic Sieve algorithm (QS),
developed 1982 by Pomerance. This effort took 8 months.

Please refer to:
\begin{list}{}{\setlength{\topsep}{-7 pt}}
   \item[] C. Pomerance: \\ {\em The quadratic sieve factoring algorithm.}
           In: G.R. Blakley / D. Chaum (Hrsg.): Proceedings of Crypto '84, LNCS 196, Springer Berlin 1995, S. 169-182
\end{list}\vskip +12 pt

In 1999 a 154-digit module (512 bit) was factored with an implementation
of the General Number Field Sieve algorithm (GNFS), developed by Buhler,
Lenstra and Pomerance. The GNFS is more efficient than QS if $n$ is longer
than about 116 decimal digits. This effort took 5 months.

Please refer to:
\begin{list}{}{\setlength{\topsep}{-7 pt}}
   \item[] J.P. Buhler / H.W. Lenstra / C. Pomerance: \\
           {\em Factoring integers with the number field sieve.} 
           In: A.K. Lenstra / H.W. Lenstra (Hrsg.): The Development of the Number Field Sieve, Lecture Notes 
           in Mathematics, Vol. 1554, Springer, Heidelberg 1993, S. 50$-$94
\end{list}\vskip +12 pt

This made evident, that a module length of 512 bit no longer prevents from
attackers.

Within the last 20 years a lot of progress has been made. Estimations about
the future development of the ability to factor RSA modules vary and depend
on some assumptions:
\begin{itemize}
\item progression in computing performance (Moore's law: every 18 month the
  computing power will double) and in grid computing 
\item development of new algorithms.
\end{itemize}
Within the last years the module bit length feasible for factorization
increased --- even without new algorithms --- by 10 bit per year. Larger
numbers require not only more time to be factored, but also more RAM
storage for the solutions matrix being used by best algorithms known today.
This need for storage grows quadratically with the length of the module
$l$.  Therefor bigger breakthroughs probably require totally new algorithms.

Within the article ``Vorz\"uge und Grenzen des RSA-Verfahrens'' Dirk Fox
prognosticates an almost linear progression with the factorization, if all
factors are included: Each year the module length feasible for
factorization increases by 20 bit on average.  Please refer: DuD, 26/2002,
pp. 84-89, F.~Bourseau / D.~Fox / C.~Thiel

The company Secorvo GmbH made a statement on the recommendation for key
length selection published by the GISA\index{GISA} (German Information
Security Agency). Chapter 2.3.1 of their statement contains a very
competent and understandable discussion of RSA security (this document
exists to my
knowledge only in German):\\
\href{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}{\texttt{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}}

\subsubsection{D.~J.\ Bernstein's paper and it's implication on the
  security of the RSA algorithm}
\label{RSABernstein} \index{Factorisation problem}

The security of the RSA algorithm is based on the empirical observation
that factoring large numbers is a hard problem. A Module $n$ (typically,
1024 bit) can be easily constructed as the product of two large primes $p$,
$q$ (typically, 500$-$600 bit each), by calculating $n=pq$. However, it is
a hard problem to extract $p$, $q$ from $n$.  Without knowing $p$ or $q$,
the private key cannot be calculated.

Thus, any progress in efficiency of factorizing large integers will effect
the security of the RSA. As a consequence, the underlying primes $p$, $q$
and, thus, the module n (1024 bit as off today) have to be increased. In
the worst case of a breakthrough in factorization, the RSA algorithm might
be compromised.

In his paper ``Circuits for integer factorization: a proposal'' (s.
\href{http://cr.yp.to/djb.html}{\texttt{http://cr.yp.to/djb.html}}),
published November 2001, D.~J.\ Bernstein addresses the problem of
factorizing large integers. Therefore, his results are of relevance from a
RSA point of view.  As a main result Bernstein claims that the
implementation of the General Number Field Sieve algorithm (GNFS) can be
improved to factor, with the same effort as before, integers with three
times more digits.

We note that the definition of \emph{effort} is a crucial point: Bernstein
claims that effort is the product of time and costs of the machine
(including the memory used). The gist of the paper lies in the fact that he
can reduce a big part of factorizing to sorting. Using Schimmler's scheme,
sorting can be optimized by massive parallel computing.  At the end of
section 3 Bernstein explains this effect: The costs of $m^2$ parallel
computers with a constant amount of memory is a constant times $m^2$.  The
costs of a computer with a single processor and memory of size $m^2$ is
also of the order of $m^2$, but with a diffrent constant factor.  With
$m^2$ processors in parallel, sorting of $m^2$ numbers (with Schimmler's
scheme) can be achieved in time $m$, while a $m^2$-memory computer needs
time of the order of $m^2$. Decreasing memory and increasing the number of
processors, one can reduce the computing time by a factor $1/m$ without
additional effort in terms of total costs.  In section 5 it is said that
massive parallel computing can also increase efficiency of factorizing
using Lenstra's elliptic-curve-method (a search algorithm has costs that
increase quadratically instead of cubically).

We note that all results achieved so far are asymptotic results. This means
that they only hold in the limit n to infinity. Unfortunately, there are no
upper bounds for the residual error (i.e. the difference between the real
and the asymptotic value) for finite n --- a problem which has already been
addressed by the author. As a consequence, one cannot conclude whether the
costs (in the sense of Bernstein) for factorizing 1024$-$2048-bit RSA modules
can be significantly reduced.

There is no doubt that Bernstein's approach is innovative. However, the
reduction of computing time under constant costs comes along with a massive
use of parallel computing --- a scenario which seems not to be realistic
yet. For example, formally 1 sec computing time on one machine and
1/1,000,000 sec time parallel computing time on 1,000,000 machines might
have same costs.  In reality, it is much harder to realize the second
situation, and Bernstein does not take into account the fixed costs, in
particular for building a network between all these computers.

Although distributed computing over a large network might help to overcome
this problem, realistic costs for data transfer has to be taken into
account --- a point which was not addressed in Bernstein's proposal.

As long as there is neither (low cost) hardware nor a distributed computing
approach (based on Bernstein's ideas), there should not be a problem for
RSA. It has to be clarified from which magnitude of n on Bernstein's method
could lead to a significant improvement (in the sense of the asymptotic
result).

Arjen Lenstra, Adi Shamir et. al. analyzed the paper of Bernstein
\cite{Lenstra2002}.  In summary they expect a factorization improvement on
how much longer the bit length of the keys could be with a factor of 1.17
(instead of factor 3 as proposed by Bernstein).

The abstract of their paper ``Analysis of Bernstein's Factorization
Circuit'' says:

``In [1], Bernstein proposed a circuit-based implementation of the matrix
step of the number field sieve factorization algorithm. We show that under
the non-standard cost function used in [1], these circuits indeed offer an
asymptotic improvement over other methods but to a lesser degree than
previously claimed: for a given cost, the new method can factor integers
that are 1.17 times larger (rather than 3.01).  We also propose an improved
circuit design based on a new mesh routing algorithm, and show that for
factorization of 1024-bit integers the matrix step can, under an optimistic
assumption about the matrix size, be completed within a day by a device
that costs a few thousand dollars.  We conclude that from a practical
standpoint, the security of RSA relies exclusively on the hardness of the
relation collection step of the number field sieve.''

\href{http://www.rsasecurity.com/}{RSA Security's} analysis of the Bernstein
paper (title {\em Has the RSA algorithm been compromised as a result of
  Bernstein's Paper?}, April, 8 2002) also concludes, that RSA is still not
compromised (s.\ 
\href{http://www.rsasecurity.com/}{\texttt{http://www.rsasecurity.com/}}).

\begin{minipage}{\textwidth}
Links:
\begin{itemize}
  \item[] \href{http://cr.yp.to/djb.html}
               {\texttt{http://cr.yp.to/djb.html}}
  \item[] \href{http://www.counterpane.com/crypto-gram-0203.html\#6}
               {\texttt{http://www.counterpane.com/crypto-gram-0203.html\#6}}
  \item[] \href{http://www.math.uic.edu}
               {\texttt{http://www.math.uic.edu}}
\end{itemize}
\end{minipage}

\vspace{4ex}
\subsubsection{Status regarding factorisation of large numbers}
\label{NoteFactorisation}

\begin{center}
\fbox{\parbox{15cm}{
    \emph{Hermann Hesse\footnotemark:}\\
    To let the possible happen, you again and again have to try the impossible.
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  German/Swiss writer and Nobel prize winner, July 2, 1877 $-$ August 9, 1962.}

The web page
\href{http://www.crypto-world.com}{\texttt{http://www.crypto-world.com}}\ 
contains an excellent overview about the factoring records of composed
integers using different methods.  The current record (as of June 2002)
obtained using the GNFS method (General Number Field Sieve) factorized a
158-digit into its both prime factors (these are build with 73 and 86
decimal digits).

This record, established on January 18, 2002 by researchers of the German
university of Bonn, got much less attention within the press than the
solution of the RSA challenge, established on August 22, 1999, when
researchers from the Netherlands factorized the 155-digit number into its
both 78-digit primes.

The task of the reserachers from Bonn was not initiated by a challenge, but
they wanted to find the last prime factors of the integer $2^{953} - 1$
(s.~ ``Wanted List'' des Cunningham Projekts
\href{http://www.cerias.purdue.edu/homes/ssw/cun/}{\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}).
The last remaining factor, called ``C158'', was known to be composite by
then, but its factors where not known:
$$
\begin{array}{c}
39505874583265144526419767800614481996020776460304936 \\
45413937605157935562652945068360972784246821953509354 \\
4305870490251995655335710209799226484977949442955603
\end{array}
$$
The factorisation resulted in the following two prime factors:
$$
\begin{array}{c}
3388495837466721394368393204672181522 \\
815830368604993048084925840555281177
\end{array}
$$
and
$$
\begin{array}{c}
1165882340667125990314837655838327081813101 \\
2258146392600439520994131344334162924536139.
\end{array}
$$

@@

\bibitem[Yan2000]{Yan2000} \index{Yan 2000} 
    Song Y. Yan, \\
    Number Theory for Computing, Springer, 2000.
\bibitem[Lenstra2002]{Lenstra2002} \index{Lenstra 2002}
    Arjen K. Lenstra, Adi Shamir, Jim Tomlinson, Eran Tromer,\\
    {\em Analysis of Bernstein’s Factorization Circuit},\\
    \href{http://www.cryptosavvy.com/mesh.pdf}{\texttt{http://www.cryptosavvy.com/mesh.pdf}}
\end{thebibliography}

@@

          \href{http://www.cryptool.com}{\texttt{http://www.cryptool.com}}
   \item Mathematica, \\
    Commercial mathematics package \\
          \href{http://www.wolfram.com}{\texttt{http://www.wolfram.com }}
   \item  LiDIA,\index{LiDIA} \\
      Extensive library containing number-theory functions and the LC interpreter \\
          \href{http://www.informatik.tu-darmstadt.de/TI/LiDIA}{\texttt{http://www.informatik.tu-darmstadt.de/TI/LiDIA}}
   \item BC, \\
     Interpreter with number-theory functions \\
          \href{http://www.maths.uq.edu.au/~krm/gnubc.html}{\texttt{http://www.maths.uq.edu.au/\~{}}krm/gnubc.html}
   \item Pari-GP, \index{Pari-GP}\\
         Excellent fast open source interpreter with number theoretical functions\\
         \href{http://www.parigp-home.de}
              {\texttt{http://www.parigp-home.de}} and
         \href{http://www.parigp-home.com}
              {\texttt{http://www.parigp-home.com}}
   \item Only after I had completed this article I did come across the
     website of Mr.\ M\"unchenbach, which interactively and 
     didactically uses elementary number theory to provide an extremely sophisticated description of the
     fundamental mathematical thought processes. It was created for a teaching project in the 11th grade 
     of the technical grammar school (unfortunately only available in German):\index{Muenchenbach} \\
         \href{http://www.djh-freeweb.de/~muenchenbach}{\texttt{http://www.djh-freeweb.de/\~{}muenchenbach}}
    \item GISA,\index{GISA} \\
          German Information Security Agency \\
          \href{http://www.bsi.bund.de}{\texttt{http://www.bsi.bund.de}} 
\end{enumerate}

@@

% Local Variables:
% TeX-master: "../script-en.tex"
% End:
